Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-25 VMCAI-revised-version/final.tex
effectivity

%%%%%
\noindent
{\bf Resiliences.}  
%
%			If the home-space contains a single element, this element is an {\em home-state}.
Given a transition system (where $S$ is the set of states) and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$ (i.e., $\post^*(X) \subseteq \pred^*(H)$).
%
A transition system is {\em resilient for $(\Bad,\Safe)$} if $(\Bad,\Safe)$ satisfies the home-space property. A transition system is {\em resilient for $\Safe$} if it is resilient for $(\Bad,\Safe)$ with $\Bad=S \setminus \Safe$. It is {\em state-resilient} if it is resilient for $(\Bad,\Safe)$ with $\Bad=\{s_0\}$.
%
%			(resp. $\post^*(s_0)$)
%                  It  is state-resilient if
%			from every state, there is a run that goes into $\Safe$.
%
%			$\mathscr{S}$ is {\em resilient for $\Safe$} if from every state, there is a run that goes into  $\Safe$.
%			$S \subseteq \pred^*(\Safe)$. 
%			this is the $(S,\Safe)$ home-space problem.
%			($\post^*(S) \subseteq \pred^*(\Safe)$).
%
% 		for $S,\Safe$: $\mathscr{S}$ is $\Safe$-resilient if $\post^*(S) \subseteq \pred^*(\Safe)$.
%		Similarly, $\mathscr{S}$ is $\Safe$-state-resilient for an initial state $s_0$ if  
%		$\post^*(s_0) \subseteq \pred^*(\Safe)$.
%		We will study three decidability questions for each type of set $\Safe$.
%		The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient 			(for a given subset of states $\Safe \subseteq S$).
%		The resilience problem could be easily generalized for two subsets $\Bad$ and $\Safe$ by asking
%		whether $\Bad \subseteq \pred^*(\Safe)$.
The {\em $k$-resilience} problem, for $k\geq0$, is to decide whether from any state is it always possible to reach $\Safe$ with a run of length smaller than $k$ (i.e., $S \subseteq \pred^{\leq k}(\Safe)$) and 
the {\em bounded resilience} problem is to decide whether there exists an $k$ such that the system is $k$-resilient.

@@

In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$ (i.e., $\mathop{\uparrow} \post^*(s)$) when $\Safe$ is upward-closed.
%
%			(called effective 
%			$\mathop{\uparrow}$ $\post^*$ basis) in the case $\Safe=\mathop{\uparrow} \Safe$ and $\Bad = \mathop{\downarrow} \Bad$.
%%%%%%%%%

@@

\noindent
{\bf Our contributions}
\begin{itemize}
\item Surprinsingly, the general undecidability statements about resilience were not known neither proved. We show that resilience and {state-resilience} (i.e. $X=\{s_0\}$) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed. 
Moreover, {state-resilience},
{bounded-state-resilience} and
{$k$-state-resilience}
are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
$\Safe$ is upward-closed. 
%			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
\item When $\Safe$ is upward-closed, the three resilience problems are decidable for both post-ideal-effective WBTS with strong (upward) compatibility and for effective pred-basis WSTS with strong (upward) compatibility. Moreover, the resilience problem is decidable for post-ideal-effective downward-WSTS with 
$\Safe$ upward-closed
%and
%the additional hypothesis that
%for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
%
%%%
%
\item We clarify the different effectiness conditions that allow decidability of the six resilience.
%
\item We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis. We show that removing the effective 
$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property. We still extend  \cite{DBLP:journals/corr/PrasadZ16} as follows: the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for post-ideal-effective WSTS with strong downward compatibility).
%
\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
%
\item We study the resilience problems for VASS and variations of VASS where most of the resilience problems are shown decidable.
\end{itemize}

>>>>

%%%%%
\noindent
{\bf Resiliences.}  
%
%			If the home-space contains a single element, this element is an {\em home-state}.
Given a transition system (where $S$ is the set of states) and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$ 
%				(i.e., $\post^*(X) \subseteq \pred^*(H)$).
%
A transition system is {\em resilient for $(\Bad,\Safe)$} if $(\Bad,\Safe)$ satisfies the home-space property. A transition system is {\em resilient for $\Safe$} if it is resilient for $(\Bad,\Safe)$ with $\Bad$ is the complementary to $ \Safe$. It is {\em state-resilient} if it is resilient for $(\Bad,\Safe)$ where $\Bad$ contains an unique state.
%
%			(resp. $\post^*(s_0)$)
%                  It  is state-resilient if
%			from every state, there is a run that goes into $\Safe$.
%
%			$\mathscr{S}$ is {\em resilient for $\Safe$} if from every state, there is a run that goes into  $\Safe$.
%			$S \subseteq \pred^*(\Safe)$. 
%			this is the $(S,\Safe)$ home-space problem.
%			($\post^*(S) \subseteq \pred^*(\Safe)$).
%
% 		for $S,\Safe$: $\mathscr{S}$ is $\Safe$-resilient if $\post^*(S) \subseteq \pred^*(\Safe)$.
%		Similarly, $\mathscr{S}$ is $\Safe$-state-resilient for an initial state $s_0$ if  
%		$\post^*(s_0) \subseteq \pred^*(\Safe)$.
%		We will study three decidability questions for each type of set $\Safe$.
%		The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient 			(for a given subset of states $\Safe \subseteq S$).
%		The resilience problem could be easily generalized for two subsets $\Bad$ and $\Safe$ by asking
%		whether $\Bad \subseteq \pred^*(\Safe)$.
The {\em $k$-resilience} problem, for $k\geq0$, is to decide whether from any state is it always possible to reach $\Safe$ with a run of length smaller than $k$ 
%
%		(i.e., $S \subseteq \pred^{\leq k}(\Safe)$)
%
 and 
the {\em bounded resilience} problem is to decide whether there exists an $k$ such that the system is $k$-resilient.

@@

In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$ 
%
%			(i.e., $\mathop{\uparrow} \post^*(s)$) 
%
when $\Safe$ is upward-closed.
%
%			(called effective 
%			$\mathop{\uparrow}$ $\post^*$ basis) in the case $\Safe=\mathop{\uparrow} \Safe$ and $\Bad = \mathop{\downarrow} \Bad$.
%%%%%%%%%

@@

\noindent
{\bf Our contributions}
\begin{itemize}
\item Surprinsingly, the general undecidability statements about resilience were not known neither proved. We show that resilience and {state-resilience} (i.e. $X=\{s_0\}$) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed. 
Moreover, {state-resilience},
{bounded-state-resilience} and
{$k$-state-resilience}
are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
$\Safe$ is upward-closed. 
%			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
\item When $\Safe$ is upward-closed, the three resilience problems are decidable for both post-ideal-effective WBTS with strong (upward) compatibility and for effective pred-basis WSTS with strong (upward) compatibility. Moreover, the resilience problem is decidable for post-ideal-effective downward-WSTS with 
$\Safe$ upward-closed
%and
%the additional hypothesis that
%for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
%
%%%
%
\item We clarify the different effectiness conditions that allow decidability of the six resilience.
%
\item We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis. We show that removing the effective 
$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property. We still extend  \cite{DBLP:journals/corr/PrasadZ16} as follows: the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for post-ideal-effective WSTS with strong downward compatibility).
%
\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$,
%
%			$\mathop{\uparrow} \post^*(s)$ 
%
for all states $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
%
\item We study the resilience problems for VASS and variations of VASS where most of the resilience problems are shown decidable.
\end{itemize}
